---
title: "Introduction to budgetr"
author: "Derek Damron"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_document:
    toc: true
    toc_float: true
    css: css/vignette.css
vignette: >
  %\VignetteIndexEntry{Introduction to budgetr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo=FALSE}
# Load libraries
library(budgetr)
library(pander)
library(knitr)
```

## What is budgetr?

*budgetr*, as you might have guessed, is an R package that helps you easily create budgets.

This introductory vignette provides a brief description of the functions provided by *budgetr* along with some simple examples on how to construct your budget.  Please refer to the function documentation for more technical information on each of these functions.

## Deconstructing a Budget

*budgetr* makes it easy to extend your recurring monthly cash flow out over time.  To do this, *budgetr* uses three primary abstractions:

1. **Items**: The individual pieces that make up your budget.
    - E.g. pay checks, rent, groceries
2. **Schedules**: A collection of **items** that detail your cash flow.
3. **Budgets**: This is a **schedule** that has been applied over a set period of time.

## A Quick Example

Let's consider a simple example to illustrate these abstractions.

Say that we have three things that constitute our monthly budgets:

1. Our monthly paycheck paid on the first day of the month
2. Our rent payment due on the first day of the month
3. Our internet bill due on the fifteenth of the month

Our first step in budgeting is to create the items themselves, which we can do with the `create_item` function:

```{r}
paycheck <- create_item( name = "Paycheck"
                       , amount = 1050
                       , day = 1
                       , recurring = TRUE
                       )
rent <- create_item( name = "Rent"
                   , amount = -800
                   , day = 1
                   , recurring = TRUE
                   )
internet <- create_item( name = "Internet"
                       , amount = -100
                       , day = 15
                       , recurring = TRUE
                       )
```

Now we need to create a schedule out of these items, which we can do with the `create_schedule` function:

```{r}
our_schedule <- create_schedule(paycheck, rent, internet)
```

Lastly (but not leastly!) we need to create our budget, which we can do with the `create_budget` function:

```{r}
our_budget <- create_budget(our_schedule)
```

Let's see what it looks like!

```{r}
our_budget
```

Looks like we're netting $150 a month, but let's plot it out to get a visual!

```{r}
plot(our_budget)
```

Check out the documentation pages for these functions for more information on their defaults and uses!  More functions and flexibility will be coming soon!
