---
title: "Introduction to budgetr"
author: "Derek Damron"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_document:
    toc: true
    toc_float: true
    css: css/vignette.css
vignette: >
  %\VignetteIndexEntry{Introduction to budgetr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo=FALSE}
# Load libraries
library(budgetr)
library(pander)
library(knitr)
```

## What is budgetr?

*budgetr*, as you might have guessed, is an R package that helps you easily create budgets.

This introductory vignette provides a brief description of the functions provided by *budgetr* along with some simple examples on how to construct your budget.  Please refer to the function documentation for more technical information on each of these functions.

## Deconstructing a Budget

*budgetr* makes it easy to extend your recurring monthly cash flow out over time.  To do this, *budgetr* uses three primary abstractions:

1. **Items**: The individual pieces that make up your budget.
    - E.g. pay checks, rent, groceries
2. **Schedules**: A collection of **items** that detail your cash flow.
3. **Budgets**: A **schedule** that has been applied over a set period of time.

## A Quick Example

Let's consider a simple example to illustrate these abstractions.

Say that you have three things that constitute your monthly budget:

1. Your monthly paycheck paid on the first day of each month
2. Your rent payment due on the fifth day of each month
3. Your internet bill due on the fifteenth of each month

Your first step in budgeting is to create the items themselves, which you can do with the `create_item` function:

```{r}
# Create a paycheck item
paycheck <- create_item( name = "Paycheck"
                       , amount = 1000
                       , day = 1
                       , recurring = TRUE
                       )
# Create a rent item
rent <- create_item( name = "Rent"
                   , amount = -500
                   , day = 5
                   , recurring = TRUE
                   )
# Create an internet item
internet <- create_item( name = "Internet"
                       , amount = -100
                       , day = 15
                       , recurring = TRUE
                       )
```

Next you need to create a schedule out of these items, which you can do with the `create_schedule` function:

```{r}
your_schedule <- create_schedule(paycheck, rent, internet)
```

Lastly (but not leastly!) you need to create your budget, which you can do with the `create_budget` function:

```{r}
your_budget <- create_budget(your_schedule, initial=1000)
```

Let's see what it looks like!

```{r}
your_budget
```

Looks like you're netting $400 a month, but let's plot it out to get a visual!

```{r}
plot(your_budget)
```

Check out the documentation pages for these functions for more information on their defaults and uses!  More functions and flexibility will be coming soon!
