---
title: "Introduction to budgetr"
author: "Derek Damron"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_document:
    toc: true
    toc_float: true
    css: css/vignette.css
vignette: >
  %\VignetteIndexEntry{Introduction to budgetr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo=FALSE, include=FALSE}
# Load libraries
library(budgetr)
library(pander)
library(knitr)
```

## What is budgetr?

*budgetr*, as you might have guessed, is an R package that helps you easily create budgets.

This introductory vignette provides a brief description of the functions provided by *budgetr* along with some simple examples on how to construct your budget.  Please refer to the function documentation for more technical information on each of these functions.

## Building a Budget

*budgetr* makes it easy to extend your recurring monthly cash flow out over time.  To do this, *budgetr* uses three primary abstractions:

1. **Items**: The individual pieces that make up your budget.
    - E.g. pay checks, rent, groceries
2. **Schedules**: A collection of **items** that detail your cash flow.
3. **Budgets**: A **schedule** that has been applied over a set period of time with a certain initial amount.

## A Quick Example

Say that you're budgeting for 2017 and you have three things that constitute your monthly budget:

1. Your monthly paycheck paid on the first day of each month
2. Your rent payment due on the fifth day of each month
3. Your internet bill due on the fifteenth of each month

Your first step in budgeting is to create the items themselves, which you can do with the `create_item` function:

```{r}
# Create a paycheck item
paycheck <- create_item( name = "Paycheck"
                       , amount = 1000
                       , day = "2017-01-01"
                       , recurring = "monthly"
                       )
# Create a rent item
rent <- create_item( name = "Rent"
                   , amount = -500
                   , day = "2017-01-05"
                   , recurring = "monthly"
                   )
# Create an internet item
internet <- create_item( name = "Internet"
                       , amount = -100
                       , day = "2017-01-05"
                       , recurring = "monthly"
                       )
```

Next you need to create a schedule out of these items, which you can do with the `create_schedule` function:

```{r}
your_schedule <- create_schedule(paycheck, rent, internet)
```

Lastly (but not leastly!) you need to create your budget, which you can do with the `create_budget` function:

```{r}
your_budget <- create_budget(your_schedule, start="2017-01-01", initial=1000)
```

Let's see what it looks like!

```{r}
your_budget
```

Looks like you're netting $400 a month, but let's plot it out to get a visual!

```{r}
plot(your_budget)
```

## A Realistic Example

In reality we have a lot more than three items in our budget.  Let's say that we *actually* have 
the following things in our budget for the next few months:

* Paychecks on the 1st and 15th of each month
* Groceries every week
* Rent on the 5th day of each month
* Internet on the 15th of each month
* Student loan on the 23rd of each month
* Car payment on the 25th of each month

First we need to create those items:

```{r, echo=FALSE}
rm(list=ls())
```

```{r}
# Paycheck items
paycheck1 <- create_item( name = "Paycheck"
                        , amount = 1000
                        , day = "2017-01-01"
                        , recurring = "monthly"
                        )
paycheck15 <- create_item( name = "Paycheck"
                         , amount = 1000
                         , day = "2017-01-15"
                         , recurring = "monthly"
                         )
# Grocery items
groceries <- create_item( name = "Groceries"
                        , amount = -100
                        , day = "2017-01-01"
                        , recurring = "weekly"
                        )
# Rent item
rent <- create_item( name = "Rent"
                   , amount = -500
                   , day = "2017-01-05"
                   , recurring = "monthly"
                   )
# Internet item
internet <- create_item( name = "Internet"
                       , amount = -100
                       , day = "2017-01-15"
                       , recurring = "monthly"
                       )
# Student loan item
student_loan <- create_item( name = "Student Loan"
                           , amount = -100
                           , day = "2017-01-23"
                           , recurring = "monthly"
                           )
# Car item
car <- create_item( name = "Car"
                  , amount = -200
                  , day = "2017-01-25"
                  , recurring = "monthly"
                  )
```

It would be a pain to have to type out all of those items, so let's use `get_items` to 
grab all of the items in our environment to create our schedule!

```{r}
your_schedule <- create_schedule(get_items())
```

We then create our budget!

```{r}
your_budget <- create_budget(your_schedule, start="2017-01-01", initial=2000)
your_budget
plot(your_budget)
```

#### Reusing in the future

We'll need to recreate our budget every so often.  There are several ways to do so depending on
your personal preference, which include but are not limited to:

* Write an R script that you can use to recreate all of your items, schedules, and budgets when needed
* Use base R functions like `save` and `saveRDS` to save a workspace with your items or to save your
individual items, schedules, and budgets

#### Updating objects

But what if our rent goes up?  It's easy to update our `rent` item with the `update_item` function.
Any arguments not specified (e.g. `name`) will be remain unchanged.

```{r}
rent <- update_item(rent, amount=-750)
```

We can then recreate our schedule (using the `get_items()` function to include all items in our current environment) and budget to see if anything substantial changed!

```{r}
your_schedule <- create_schedule(get_items())
your_budget <- create_budget(your_schedule, start="2017-01-01", initial=2000)
your_budget
plot(your_budget)
```

There are also functions to quickly update your existing schedules and budgets (e.g. for fast scenario
testing):

* `update_schedule` lets you quickly add and/or remove items from a particular schedule.  For example, say that you forgot to include the deposit for a late-received Christmas present: 

```{r}
xmas_present <- create_item( name = "Xmas Present"
                           , amount = 200
                           , day = "2017-01-01"
                           )
your_schedule_updated <- update_schedule(your_schedule, add=xmas_present)
```

* `update_budget` lets you quickly change the start date, the end date, and/or the initial amount.  For example, say that you wanted to update `your_budget` to apply through the end of 2017:

```{r}
your_budget_2017 <- update_budget(your_budget, end="2017-12-31")
plot(your_budget_2017)
```
